---
title: "Pizza Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pizza_plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(soccr)
library(tidyverse)
library(devtools)
library(gt)
load_all()
```

The function `plot_radar_man` is meant to be a simple way to create pizza plots given a simple data frame. It currently requires the data frame to have columns named specifically as such: `statistic`, `value`, `percentile`, `category`, and `color.` The plot will follow the order of the rows in the data frame.

```{r data}
scaled_data <- data.frame(player = rep("Alexia Putellas", 12),
                          statistic = c("Non-Penalty Goals", "Assists", "Shot-Creating Actions",
                          "Pass Completion Perc", "Progressive Passes", "Progressive Carries",
                          "Tackles", "Interceptions", "Aerials Won",
                          "Fouls Drawn","Fouls Committed", "Offsides"),
                          value = c(.69, 0.1, 6.2, 80, 8.1, 2.0, 0.8, 0.4, 0.8, 0.8, 0.7, 0.7),
                          percentile = c(85, 48, 99, 99, 99, 70, 42, 85, 21, 42, 73, 61),
                          category = rep(c("Attack", "Possession", "Defending", "Misc"), each = 3),
                          color = rep(c("#008000", "#0000FF", "#FF0000", "#FFA500") ,each = 3))

scaled_data
```

```{r plot, fig.width = 6, fig.height = 6, fig.align = "center"}
p <- plot_radar_man(scaled_data,
                    title = scaled_data$player[1],
                    subtitle = paste("All Matches for 2023 | Compared to FW/AM",
                                     "All Data per 90",
                                     sep = "\n"),
                    caption = "Data: FBref.com | By: Sean Steele",
                    pos_group = "FW/AM",
                    theme_color = "light",
                    title_loc = "left")
p
```

Depending on your output, you may need to adjust the size of the plot. For instance, I set my rmarkdown chunk to `{r plot, fig.width = 6, fig.height = 6, fig.align = "center"}`, but you can see that the plot tends to get squished. In quarto, I have found the following settings render the plot nicely:

````markdown
---
format:
  html:
    page-layout: full
---

`r ''````{r plot}
#| fig.width: 10
#| fig.height: 10
#| fig.align: "center"

plot_radar_man(...)
```
````

Alternatively, you can save the plot as a png and then render it in your document.

```{r save}
ggsave("putellas.png", p, width = 2700, height = 3000, units = "px")
```

Once you have save your image, I have found that the code chunk `{r plot, out.width = "100%", fig.align = "center"}` renders the image nicely for html output via quarto or rmarkdown (utilized in the below plots). You can play with the width and height settings to add more padding or bring the image closer to the text. Because the text categories are not always the same length, and dynamically created, you may need to adjust the width of the plot to get the text to fit nicely.

Thomas Mock provides a clean way of add an image/logo to a plot. I have included it here with `plot_add_logo`.

```{r add-logo, out.width = "100%", fig.align = "center"}
logo <-"https://upload.wikimedia.org/wikipedia/en/thumb/4/47/FC_Barcelona_%28crest%29.svg/142px-FC_Barcelona_%28crest%29.svg.png?20120407190948"

plot <- "putellas.png"

plot_add_logo(plot, logo, logo_position = "top right", logo_scale = 12, logo_padding = 0.01)

```

`logo` can be either a html link or a local file.

My pizza plot function has been updated to allow the quick creation of a pizza plot with a logo. The logo will be placed in the top right corner by default. Use the above code if you want more control over the placement.

```{r plot-logo, out.width = "100%", fig.align = "center"}
plot_radar_man(scaled_data,
               title = scaled_data$player[1],
               subtitle = paste("All Matches for 2023 | Compared to FWDs",
                                "All Data per 90", 
                                sep = "\n"),
               caption = "Data: FBref.com | By: Sean Steele",
               pos_group = "FW/AM",
               theme_color = "light",
               logo_path = "https://upload.wikimedia.org/wikipedia/en/thumb/4/47/FC_Barcelona_%28crest%29.svg/142px-FC_Barcelona_%28crest%29.svg.png?20120407190948")
```
